---
title: "COVID-19_DATA Final Project"
author: "Ram"
date: "2024-12-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(lubridate)

path_in <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/"
file_names<-c("time_series_covid19_confirmed_global.csv","time_series_covid19_deaths_global.csv","time_series_covid19_confirmed_US.csv","time_series_covid19_deaths_US.csv")
urls <- str_c(path_in,file_names)
global_cases<-read_csv(urls[1])
global_deaths<-read_csv(urls[2])
US_cases<-read_csv(urls[3])
US_deaths<-read_csv(urls[4])

global_cases <- global_cases %>%
  pivot_longer(cols = -c('Province/State','Country/Region','Lat','Long'),
    names_to = "date",
    values_to = "cases") %>%
  select(-c('Lat','Long'))

global_deaths <- global_deaths %>%
  pivot_longer(cols = -c('Province/State','Country/Region','Lat','Long'),
    names_to = "date",
    values_to = "deaths") %>%
  select(-c('Lat','Long'))

global_joined <- global_cases %>%
full_join(global_deaths) %>%
rename(Country_Region = 'Country/Region', Province_State = 'Province/State') %>%
mutate(date = mdy(date))

plot_belguim_zoom_all <- global_joined %>% filter(`Country_Region` == "Belgium")
plot_france_zoom_all <- global_joined %>% filter(`Country_Region` == "France")
plot_spain_zoom_all <- global_joined %>% filter(`Country_Region` == "Spain")

plot_belguim <- global_joined %>% filter(`Country_Region` == "Belgium" & date > '2021-09-30' & date < '2022-02-01')
plot_france <- global_joined %>% filter(`Country_Region` == "France" & date > '2021-09-30' & date < '2022-02-01')
plot_spain <- global_joined %>% filter(`Country_Region` == "Spain" & date > '2021-09-30' & date < '2022-02-01')

plot_netherlands <- global_joined %>% filter(`Country_Region` == "Netherlands" & date > '2021-09-30' & date < '2022-02-01')

```



In this work, at first, I decided to get an over all look over Belgium, France and Spain\
\
I marked number-of-cases in BLUE and number-of-deaths in RED\
\
```{r ZOOM_ALL, echo=FALSE}
plot_belguim_zoom_all  %>% ggplot() + geom_point(aes(x = date,y = cases/100),color = "blue") + geom_point(aes(x=date,y = deaths),color = "red")

plot_france_zoom_all  %>% ggplot() + geom_point(aes(x = date,y = cases/100),color = "blue") + geom_point(aes(x=date,y = deaths),color = "red")

plot_spain_zoom_all  %>% ggplot() + geom_point(aes(x = date,y = cases/100),color = "blue") + geom_point(aes(x=date,y = deaths),color = "red")

```
\

\
In all countries I noticed a ramp up in the cases right near the end of 2021\
\
So, I zoomed in between October and February for each country\
\
BULGIUM\
\
```{r BELGUIM, echo=FALSE}
plot_belguim  %>% ggplot() + geom_point(aes(x = date,y = cases/100),color = "blue") + geom_point(aes(x=date,y = deaths),color = "red")
```
\
The ramp up start in middle of October\
\
FRANCE\
\
```{r FRANCE, echo=FALSE}
plot_france  %>% ggplot() + geom_point(aes(x = date,y = cases/100),color = "blue") + geom_point(aes(x=date,y = deaths),color = "red")
```
\
The ramp up start in the beginning of December\
\
SPAIN\
\
```{r SPAIN, echo=FALSE}
plot_spain  %>% ggplot() + geom_point(aes(x = date,y = cases/100),color = "blue") + geom_point(aes(x=date,y = deaths),color = "red")
```
\
The ramp up start in the middle of November\
\
\
I wanted to find out if near by countries will have the same ramp up\
I did the same analyze for the Netherlands and expected to find a ramp up as well\
\
NETHERLANDS\
\
```{r Netherlands, echo=FALSE}
plot_netherlands  %>% ggplot() + geom_point(aes(x = date,y = cases/100),color = "blue") + geom_point(aes(x=date,y = deaths),color = "red")
```
\
I found that the ramp up start in the middle of October\
\

